---
title: "What makes the playlist Borderless borderless?"
author: "Rosie Zheng"
date: "8-3-2021"
output: 
    flexdashboard::flex_dashboard:
        storyboard: true
        theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(spotifyr)
library(tidyverse)
library(dplyr)
library(hrbrthemes)
library(viridis)
library(ggplot2)
library(flexdashboard)
library(plotly)
library(compmus)
```

```{r}
borderless <- get_playlist_audio_features("", "37i9dQZF1DWZJmo7mlltU6")
pop <- get_playlist_audio_features("", "37i9dQZF1DXcBWIGoYBM5M")
hiphop <- get_playlist_audio_features("", "37i9dQZF1DX0XUsuxWHRQd")
dance <- get_playlist_audio_features("", "37i9dQZF1DXaXB8fQg7xif")
rock <- get_playlist_audio_features("", "37i9dQZF1DWXRqgorJj26U")
indie <- get_playlist_audio_features("", "37i9dQZF1DX2Nc3B70tvx0")

playlists_portfolio <-
  bind_rows(
    borderless %>% mutate(category = "Borderless"),
    pop %>% mutate(category = "Pop"),
    hiphop %>% mutate(category = "Hip-Hop"),
    dance %>% mutate(category = "Dance/Electronic"),
    rock %>% mutate(category = "Rock"),
    indie %>% mutate(category = "Indie")
  )

# Sample size
sample_size_portfolio = playlists_portfolio %>% group_by(playlist_name) %>% summarize(num=n())
```

### Tempo of Borderless songs in comparison to the other genres

```{r}
# Plot
tempo <- playlists_portfolio %>%
  left_join(sample_size_portfolio) %>%
  mutate(myaxis = paste0(playlist_name, "\n", "n=", num)) %>%
  ggplot( aes(x=myaxis, y=tempo, fill=playlist_name)) +
    geom_violin(width=1, alpha=0.5) +
    geom_boxplot(width=0.1, color="grey", alpha=0.8) +
    scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=15)
    ) +
    ggtitle("Tempo") +
    xlab("Playlist") +
  ylab("BPM")
ggplotly(tempo)
```

***

This violin plot shows the distribution of tempo across the different playlists. It also includes boxplots to show the median and quartiles. The small black dots represent the outliers.

The median tempo for the Borderless playlist lays at 115 BPM. This median comes closest to the Indie playlist and is mostly different from the Dance playlist. Furthermore, we observe that the distribution for tempo is wide-spread for the Borderless playlist and is similar to the distributions of the Rock, Pop, and Indie playlist. 

### How does the fastest Borderless song and outlier "L'amor fa calor" look like in a tempogram?

```{r}
calor <- get_tidy_audio_analysis("2xhQ9obJQFjfjCdCugCY5D")
calor %>%
  tempogram(window_size = 8, hop_size = 1, cyclic = FALSE) %>%
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()
  
calor %>%
  tempogram(window_size = 8, hop_size = 1, cyclic = TRUE) %>%
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()
```

***

These tempograms show the tempo analysis from the Borderless song "Abandoned". Both temporgrams are Fourier-based and therefore stronly tend to pick up on tempo harmonics. Sometimes it can be more informative to wrap this into a cyclic tempogram. The right tempogram shows such a cyclic tempogram. 

Regarding the violin plot for Borderless from the previous tab, songs with a tempo above 172 are considered outliers.The greatest outlier is the song with a tempo value of 185.975 and belongs to the song "Abandoned" from artist Kit Sebastian. Therefore, these tempograms will probably look different from the rest of the Borderless songs.

Listening to "Abandoned", it does not necessarily sound that there are a lot beats per minute. Also when looking at the cyclic tempogram (right), it looks like BPM is constantly around 95. For the other tempogram (left), it looks like there are two main tempos which play throughout the whole song: 200 and 400. Since both the ride symbal and the toms are constantly playing at the same time during this song, it could be the case that Spotify recognizes both instruments as beats. 

### Key distribution over the different genres. Do the Borderless keys differ from the keys in other genres?

```{r}
playlists_portfolio$key <- as.factor(playlists_portfolio$key)
keys <- ggplot(playlists_portfolio, aes(x=playlist_name, fill=key)) +
  geom_bar(position = 'fill', width=0.5) + ylab("Proportion") + xlab('Playlist name')
keys + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
  labs(title="Key distribution", 
       subtitle="Keys across the different music genres",
       xlab = 'Playlist name') 
```

***

This histogram shows the distribution of keys across the different playlists. 

**For feedback**: What should be noted is that the playlists do not have the same song count. Therefore, the heights of the bars would differ when I would do normal count but now I have taken the proportion. Or shall I delete songs from playlists manually in R in order to have equal sized playlists? Love to hear some feedback on this.  

Comparing Borderless with the other genres, it seems that Ultimate Indie shows the most similar distribution of keys. However, some keys (1, 5, 7, 11) appear more in Borderless than in Ultimate Indie. It is remarkable that key 2 is the most apparent one in RapCaviar, while this is not the case in all other playlists. It appears that the keys are generally well-distributed in the Borderless playlist. Only keys 6, 8, 10 have apparent smaller numbers but the rest seems quite evenly divided. A striking founding is that key 3 is not present in the Borderless playlist at all. This key is also rare in the other genres but is present at least. This could be one of the reasons why this playlist is called Borderless. 

### Loudness of Borderless songs in comparison to the other genres

```{r}
# Plot
loudness <- playlists_portfolio %>%
  left_join(sample_size_portfolio) %>%
  mutate(myaxis = paste0(playlist_name, "\n", "n=", num)) %>%
  ggplot( aes(x=myaxis, y=loudness, fill=playlist_name)) +
    geom_violin(width=1, alpha=0.5) +
    geom_boxplot(width=0.1, color="grey", alpha=0.8) +
    scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=15)
    ) +
    ggtitle("Loudness") +
    xlab("Playlist") +
  ylab("dB")
ggplotly(loudness)
```
***

This violin plot shows the distribution of loudness across the different playlists. It also includes boxplots to show the median and quartiles. The small black dots represent the outliers.

The median of Borderless lays closest to the median of Today's Top Hits, while the quartiles look most similar to the Ultimate Indie playlist. Also the shape of the violin of Borderless is most similar to the one of Ultimate Indie. However, it appears that Ultimate Indie includes a few silent songs while Bordeless does not have this. 

### The Borderless song Stargazing - Still Woozy compared with the pop song Heat Waves - Glass Animals.

```{r chroma}
stargazing <-
  get_tidy_audio_analysis("0VF7YLIxSQKyNiFL3X6MmN") %>%
  select(segments) %>%
  unnest(segments) %>%
  select(start, duration, pitches)
heatwaves <-
  get_tidy_audio_analysis("3USxtqRwSYz57Ewm6wWRMp") %>%
  select(segments) %>%
  unnest(segments) %>%
  select(start, duration, pitches)
bord_dist <-
  compmus_long_distance(
    stargazing %>% mutate(pitches = map(pitches, compmus_normalise, "manhattan")),
    heatwaves %>% mutate(pitches = map(pitches, compmus_normalise, "manhattan")),
    feature = pitches,
    method = "aitchison"
  )
```

```{r tallis-plot}
pop_bord <-
  bord_dist %>%
  mutate(
    stargazing = xstart + xduration / 2,
    heatwaves = ystart + yduration / 2
  ) %>%
  ggplot(
    aes(
      x = stargazing,
      y = heatwaves,
      fill = d
    )
  ) +
  geom_tile(aes(width = xduration, height = yduration)) +
  coord_fixed() +
  scale_x_continuous(
    breaks = c(0, 60, 105, 150, 185, 220, 280, 327),
    labels =
      c(
        "",
        "",
        "",
        "",
        "",
        "",
        "",
        ""
      ),
  ) +
  scale_y_continuous(
    breaks = c(0, 45, 80, 120, 145, 185, 240, 287),
    labels =
      c(
        "",
        "",
        "",
        "",
        "",
        "",
        "",
        ""
      ),
  ) +
  scale_fill_viridis_c(option = "E", guide = "none") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  labs(x = "Stargazing - Still Woozy", y = "Heat Waves - Glass Animals")
pop_bord
```

***

This visualisation of Stargazing - Still Woozy (Borderless) and Heat Waves - Glass Animals (Pop) between chroma features to show how the two songs within these genres align with one another. 

For the first four parts, the relationship between the songs is consistent: the song from Still Woozy is somewhat more slowly than the one from Glass Animals. For the fifth part, Still Woozy is singing faster than Glass Animals. However, at the beginning of the sixth part, Still Woozy return to his regular tempo relationship with Glass Animals.

Note (will be deleted later but more for feedback): in some or other way, this chromogram looks very much the same as the one in Ashley's portfolio but I changed the songs. How is this possible? These songs are very different than the ones from Ashley. Also, when I knit this, I can see that there is no observable blue diagonal line in the portfolio while there is one when I run this in R... 

### The feel&sound from ‘Rocky - Still Woozy' [chroma and timbre features]

```{r rocky}
rocky <-
  get_tidy_audio_analysis("5uDohC8URXwbGiHAMuKGoa") %>%
  compmus_align(bars, segments) %>%
  select(bars) %>%
  unnest(bars) %>%
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "acentre", norm = "manhattan"
      )
  ) %>%
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "mean"
      )
  )
bind_rows(
  rocky %>% 
    compmus_self_similarity(pitches, "aitchison") %>% 
    mutate(d = d / max(d), type = "Chroma"),
  rocky %>% 
    compmus_self_similarity(timbre, "euclidean") %>% 
    mutate(d = d / max(d), type = "Timbre")
) %>%
  mutate() %>% 
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  facet_wrap(~type) +
  scale_fill_viridis_c(option = "E", guide = "none") +
  theme_classic() + 
  labs(x = "", y = "")
```

***

The two self-similarity matrices are summarised at the bar level but have axes in seconds. They illustrate pitch- and timbre-based self-similarity within singer-songwriter Still Woozy's recently released ‘Rocky’ (19 February, 2021). Studying both matrices helps us to understand the structure of the song. It is remarkable that both matrices are very similar. It seems that only the first seconds of the song have a clearly different timbre than chroma. Furthermore, it seems like there can be no clear structure observed from the matrices while in the song itself, you can clearly hear the the difference between the verses and choruses. The song does not have a fade-out which is observable from these matrices since te pattern at the end is not different from the rest of the matrices. 


### Introduction

Spotify is the gate to many music genres of all time around the whole world. When we click on “Browse”, there is a tab called “Genres and Moods”, including playlists for the genres Pop, Hip-Hop, Dance/Electronic, Indie, Rock, Soul, R&B, Latin, Country, Folk, Jazz, Blues, Metal, Classical, K-pop, Funk, and Punk. The playlist that cannot be found in any of these genres is “Borderless” and has the following description: “Beyond genre, beyond language, beyond borders”. However, it is not clear what musical components makes this playlist so-called borderless. Are the 98 songs in this playlist really that different from our common-known genres and why exactly? Has the time come to produce “borderless” music?

In this project, we will compare the Borderless playlist with playlists from the five most popular genres nowadays on Spotify: Pop, Hip Hop, Dance/Electronic, Indie, and Rock. The playlists are selected based on highest follower number for that genre. This lead to the following selection: Today's Top Hits (Pop), RapCaviar (Hip-Hop), Dance Party (Dance/Electronic), Ultimate Indie (Indie), and Rock Classics (Rock). We expect that the musical components which belong to a specific genre will not be recognised in the Borderless playlist.

Regarding that the selection of the playlists is only based on number of followers, a limitation could be that the chosen playlists do not fully represent the whole genres, but only the opinion of the big public. However, each playlist will have at least one classic song for that genre included. We can dedicate some detailed analysis on these song and compare them to the songs from the Borderless playlist.


### Placement of Borderless songs on the happy scale (energy + valence) in comparison to the other genres

```{r}
happiness <- playlists_portfolio %>%                   
  mutate(
    mode = ifelse(mode == 0, "Minor", "Major")
  ) %>%
  ggplot(                    
    aes(
      x = valence,
      y = energy,
      size = loudness,
      color = mode
    )
  ) +
  geom_point(alpha=0.7) + 
  geom_rug(size = 0.1) +     
  geom_text(                 
    aes(
      x = valence,
      y = energy,
      label = label
    ),
    data = 
      tibble(
        label = c("i'll find my way to you tomorrow", "Lonely (with benny blanco)", "	
Walk On the Wild Side", "Frequency", "Pressure (feat. Tove Lo)", "Moonwalking in Calabasas (feat. Blueface) - Remix"),
        category = c("Borderless", "Pop", "Rock", "Indie", "Dance/Electronic", "Hip-Hop"),
        valence = c(0.6610, 0.0927, 0.385, 0.2140, 0.3570, 0.7450),
        energy = c(0.294, 0.239, 0.318, 0.200, 0.558, 0.341)
      ),
    colour = "black",        
    size = 2,                 
    hjust = "left",          
    vjust = "bottom",         
    nudge_x = -0.05,          
    nudge_y = 0.02            
  ) +
  facet_wrap(~category) +   
  scale_x_continuous(         
    limits = c(0, 1),
    breaks = c(0, 0.25, 0.5,0.75, 1), 
    minor_breaks = NULL       
  ) +
  scale_y_continuous(         
    limits = c(0, 1),
    breaks = c(0, 0.25, 0.5, 0.75, 1),
    minor_breaks = NULL
  ) +
  scale_colour_brewer(     
    type = "qual",            
    palette = "Pastel2"       
  ) +
  scale_size_continuous(     
    trans = "exp",           
    guide = "none"         
  ) +
  theme_light() +             
  labs(                      
    x = "Valence",
    y = "Energy",
    colour = "Mode"
  )

ggplotly(happiness)
```

***

Regarding these graphs, the Borderless playlist is widely distributed over valence and has energy levels of 0.25 and up. It seems that the higher the energy and valence, the louder the music. There are slighly more songs with major modalities than with minor ones. When comparing the Borderless playlist with the other genres, we observe that energy levels are similar to Hip-Hop, Pop, and Indie, valence levels are similar to Rock, loudness levels are similar to Pop and Hip-Hop, and mode is similar to Pop. 

### Tempo of Borderless songs in comparison to the other genres

```{r}
# Sample size
sample_size_portfolio = playlists_portfolio %>% group_by(playlist_name) %>% summarize(num=n())

# Plot
playlists_portfolio %>%
  left_join(sample_size_portfolio) %>%
  mutate(myaxis = paste0(playlist_name, "\n", "n=", num)) %>%
  ggplot( aes(x=myaxis, y=tempo, fill=playlist_name)) +
    geom_violin(width=1, alpha=0.5) +
    geom_boxplot(width=0.1, color="grey", alpha=0.8) +
    scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=15)
    ) +
    ggtitle("Tempo") +
    xlab("Playlist") +
  ylab("BPM")

```

***

The mean tempo for the Borderless playlist lays around 120 BPM. This mean comes closest to the Indie playlist and is mostly different from the Dance playlist. Furthermore, we observe that the distribution for tempo is wide-spread for the Borderless playlist and is similar to the distributions of the Rock, Pop, and Indie playlist. 

### Energy of Borderless songs in comparison to the other genres

```{r}
# Plot
playlists_portfolio %>%
  left_join(sample_size_portfolio) %>%
  mutate(myaxis = paste0(playlist_name, "\n", "n=", num)) %>%
  ggplot( aes(x=myaxis, y=energy, fill=playlist_name)) +
    geom_violin(width=1, alpha=0.5) +
    geom_boxplot(width=0.1, color="grey", alpha=0.8) +
    scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=15)
    ) +
    ggtitle("Energy") +
    xlab("Playlist") +
  ylab("Energy Level")

```

***

In the first graph, we observed that energy levels for Borderless were similar to Hip-Hop, Pop, and Indie. Here we can observe that the violin shape is mostly the samefor Pop and indie. Rap, however, shares the same boxplot.

### Loudness of Borderless songs in comparison to the other genres

```{r}
# Plot
playlists_portfolio %>%
  left_join(sample_size_portfolio) %>%
  mutate(myaxis = paste0(playlist_name, "\n", "n=", num)) %>%
  ggplot( aes(x=myaxis, y=loudness, fill=playlist_name)) +
    geom_violin(width=1, alpha=0.5) +
    geom_boxplot(width=0.1, color="grey", alpha=0.8) +
    scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=15)
    ) +
    ggtitle("Loudness") +
    xlab("Playlist") +
  ylab("dB")

```
### Danceability of Borderless songs in comparison to the other genres

```{r}
# Plot
playlists_portfolio %>%
  left_join(sample_size_portfolio) %>%
  mutate(myaxis = paste0(playlist_name, "\n", "n=", num)) %>%
  ggplot( aes(x=myaxis, y=danceability, fill=playlist_name)) +
    geom_violin(width=1, alpha=0.5) +
    geom_boxplot(width=0.1, color="grey", alpha=0.8) +
    scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=15)
    ) +
    ggtitle("Danceability") +
    xlab("Playlist") +
  ylab("Danceability Level")

```

***

Danceability for Borderless is mostly similar to Indie. 

### Speechiness of Borderless songs in comparison to the other genres

```{r}
# Plot
playlists_portfolio %>%
  left_join(sample_size_portfolio) %>%
  mutate(myaxis = paste0(playlist_name, "\n", "n=", num)) %>%
  ggplot( aes(x=myaxis, y=speechiness, fill=playlist_name)) +
    geom_violin(width=1, alpha=0.5) +
    geom_boxplot(width=0.1, color="grey", alpha=0.8) +
    scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=15)
    ) +
    ggtitle("Speechiness") +
    xlab("Playlist") +
  ylab("Speechiness Level")

```

***

Speechiness is low for every genre except Hip-Hop which is logical. Borderless comes closest to Rock and Indie. 

### Acousticness of Borderless songs in comparison to the other genres

```{r}
# Plot
playlists_portfolio %>%
  left_join(sample_size_portfolio) %>%
  mutate(myaxis = paste0(playlist_name, "\n", "n=", num)) %>%
  ggplot( aes(x=myaxis, y=acousticness, fill=playlist_name)) +
    geom_violin(width=1, alpha=0.5) +
    geom_boxplot(width=0.1, color="grey", alpha=0.8) +
    scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=15)
    ) +
    ggtitle("Acousticness") +
    xlab("Playlist") +
  ylab("Acousticness Level")

```

***

The distribution for acousticness is almost the same for Rock. It is completely different from Dance.

### Instrumentalness of Borderless songs in comparison to the other genres

```{r}
# Plot
playlists_portfolio %>%
  left_join(sample_size_portfolio) %>%
  mutate(myaxis = paste0(playlist_name, "\n", "n=", num)) %>%
  ggplot( aes(x=myaxis, y=instrumentalness, fill=playlist_name)) +
    geom_violin(width=1, alpha=0.5) +
    geom_boxplot(width=0.1, color="grey", alpha=0.8) +
    scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=15)
    ) +
    ggtitle("Instrumentalness") +
    xlab("Playlist") +
  ylab("Instrunemntalness Level")

```

***

It seems that instrumentalness cannot be visualized very well with a violin plot. 

### Liveness of Borderless songs in comparison to the other genres

```{r}
# Plot
playlists_portfolio %>%
  left_join(sample_size_portfolio) %>%
  mutate(myaxis = paste0(playlist_name, "\n", "n=", num)) %>%
  ggplot( aes(x=myaxis, y=liveness, fill=playlist_name)) +
    geom_violin(width=1, alpha=0.5) +
    geom_boxplot(width=0.1, color="grey", alpha=0.8) +
    scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=15)
    ) +
    ggtitle("Liveness") +
    xlab("Playlist") +
  ylab("Liveness Level")

```

***

The distribution for liveness is mostly similar to Indie but is more smeared out for higher values for Borderless. 

### Valence of Borderless songs in comparison to the other genres

```{r}
# Plot
playlists_portfolio %>%
  left_join(sample_size_portfolio) %>%
  mutate(myaxis = paste0(playlist_name, "\n", "n=", num)) %>%
  ggplot( aes(x=myaxis, y=valence, fill=playlist_name)) +
    geom_violin(width=1, alpha=0.5) +
    geom_boxplot(width=0.1, color="grey", alpha=0.8) +
    scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=15)
    ) +
    ggtitle("Valence") +
    xlab("Playlist") +
  ylab("Valence Level")
```

***

Valence would be most similar to Rock. However, there seem to be more lower valences in the Borderless playlist than in the Rock playlist.